---
title: "Final project"
author: "David"
date: "2023-05-02"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message= FALSE,
                       warning =FALSE,
                       fig.path= "../output/")
```

# Where did the Data base come from?
  The data used on theis project came from my research on "Exploring Mitochondrial DNA Variation among Ancient Indigenous American Populations". The data was obtained by UC Santa Cruz researcher collaborators, they had ground down tissue to sequence DNA. Which allowed me to do bioinformatic methods. these methods were done in the terminal using SAMtools; which included creating an index, subset, trimming, and consensus on the DNA samples to get final mtDNA samples. Then use Haplogrep(an online tool to identify haplogroups for individual samples) to get haplogroup data.




# Library
```{r library}
library(tidyverse)
library(here)
library(ggthemes)
library(dplyr)
library(tidytext)
library(ggplot2)
library(cowplot)
```

# Data import
```{r}
mtDNA <- read_csv(here("data", "filtered_haploid.csv"))%>%
  mutate(new = str_sub(`Haplogroup(trim)`,start=1, end= 2))%>% #filtering row names for haplogroup for trim
  mutate(region= str_sub(SampleID, start = 5, end= 6)) # filter sample ID name to get region
```

# Data dictionary
```{r}
data_dictionary <- read_csv(here("data", "data_dictionary.csv"))

glimpse(data_dictionary)
```

# First plot
  This graph shows the distribution of which haplogroups were found within the mtDNA samples

```{r}
#filtered data
new_mtDNA <- read_csv(here("data", "filtered_haploid.csv"))%>%
  mutate(new = str_sub(`Haplogroup(trim)`,start=1, end= 2))%>% #filtering row names for haplogroup for trim
  mutate(region= str_sub(SampleID, start = 5, end= 6))%>% # filter sample ID name to get region
  group_by(new)

#script for second plot 
new_mtDNA%>%
  ggplot(aes(y= new))+
  geom_bar(fill= "navy blue")+
  geom_text(stat = "count", aes(label = ..count..),color = "white", position = position_stack(vjust = 0.5))+## this line isnt working
  theme_classic()+
  labs(x= "Amount of haplogroups",
       y= "Haplogroups",
       title= "Amount of haplogroups groups counted from sample")
```

# The locations where samples were taken from:
  Samples were taken from different regions within Columbia and Peru.

- CL: Colombia  
- NC: Northern Coast  
- SC: Southern Coast  
- NH: Northern Highlands  
- SH: Southern Highlands  
- TR: Amazonian ancestry that were buried in SH  


# second plot
  This stack plot shows the distribution of which haplogroup was found within each location. Within the stack plot, the amount of haplogroups counted for each region is labeled on the bar. 

```{r pressure, echo=FALSE}
#data for plot 1
maybe_mtDNA <- mtDNA%>% ##test for stacking plot
  group_by(region)%>%
  count(new)
  
# plot script
distribution <- 
maybe_mtDNA %>%
  ggplot(aes( y=n, x=region, fill=new)) +
  geom_bar(position="fill", stat="identity")+
  theme_calc(base_size = 10 ,base_family = "sans")+
  labs(x= "Region where sample was found",
       y= "Percentage of haplogroups",
       title = "mtDNA haplogroup distribution across regions (percentage)") +
  scale_fill_discrete(name = "Haplogroup") 

ggdraw(distribution)+
  draw_label("2", x = 0.13, y = 0.7, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in CL for A2 haplogroups
    draw_label("2", x = 0.13, y = 0.3, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in CL for B2 haplogroups
    draw_label("1", x = 0.235, y = 0.862, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NC for A2 haplogroups
  draw_label("3", x = 0.235, y = 0.75, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NC for B2 haplogroups
  draw_label("6", x = 0.235, y = 0.5, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NC for C haplogroups
  draw_label("4", x = 0.235, y = 0.23, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NC for D1 haplogroups
  draw_label("4", x = 0.345, y = 0.8, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NH for A2 haplogroups
   draw_label("3", x = 0.345, y = 0.66, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NH for B2 haplogroups
   draw_label("9", x = 0.345, y = 0.4, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NH for C haplogroups
   draw_label("3", x = 0.345, y = 0.175, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in NH for D1 haplogroups
   draw_label("3", x = 0.455, y = 0.745, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in SC for A2 haplogroups
  draw_label("2", x = 0.455, y = 0.5, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in SC for B2 haplogroups
  draw_label("3", x = 0.455, y = 0.24, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in SC for C haplogroups
  draw_label("4", x = 0.56, y = 0.82, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in SH for A2 haplogroups
  draw_label("11", x = 0.55, y = 0.55, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in SH for B2 haplogroups
  draw_label("3", x = 0.56, y = 0.274, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in SH for C haplogroups
  draw_label("3", x = 0.56, y = 0.17, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in SH for D1 haplogroups
  draw_label("1", x = 0.67, y = 0.49, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in TI for A2 haplogroups
  draw_label("1", x = 0.77, y = 0.77, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+ #label for number of samples in TR for B2 haplogroups
  draw_label("2", x = 0.77, y = 0.39, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black") #label for number of samples in TR for C haplogroups

```

# Third Plot
The donut graph gives the viewers a different perception on amount of haplogroups found from samples used.
```{r}
### testing third plot, I HATE THIS!!!!!!!!!!!!!!
testing_mtDNAnew <- read_csv(here("data", "filtered_haploid.csv"))%>%
  mutate(new = str_sub(`Haplogroup(trim)`,start=1, end= 2))%>% #filtering row names for haplogroup for trim
  mutate(region= str_sub(SampleID, start = 5, end= 6))%>% # filter sample ID name to get region
  group_by(new)

test_mtDNA <- data.frame(region= c("A2", "B2", "C", "D1"),
  amount= c(14,23,23,10)
  )

test_mtDNA$fraction = test_mtDNA$amount / sum(test_mtDNA$amount)# Compute percentages

test_mtDNA$ymax = cumsum(test_mtDNA$fraction) # Compute the cumulative percentages (top of each rectangle)

test_mtDNA$ymin = c(0, head(test_mtDNA$ymax, n=-1)) # Compute the bottom of each rectangle

# Make the plot

donut <- 
ggplot(test_mtDNA, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=2, fill=region)) + #the fill=category makes my code stop
  geom_rect() +
  coord_polar(theta="y") + # Try to remove that to understand how the chart is built initially
  xlim(c(-3, 4))+ # Try to remove that to see how to make a pie chart
  theme(axis.text.x= element_blank(),
        axis.text.y= element_blank(),
        axis.line.x= element_blank(), 
        axis.line.y= element_blank(),
        axis.ticks.y = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  scale_fill_discrete(name = "Haplogroup") 

ggdraw(donut) + #uses plot to layer on top
draw_label("23", x = 0.59, y = 0.3, hjust = 0, vjust = 0,
           fontfamily = "", fontface = "plain", color = "black", size = 14,
           colour = "black")+
  draw_label("14", x = 0.52, y = 0.77, hjust = 0, vjust = 0,
             fontfamily = "", fontface = "plain", color = "black", size = 14,
             colour = "black")+
  draw_label("23", x = 0.19, y = 0.4, hjust = 0, vjust = 0,
             fontfamily = "", fontface = "plain", color = "black", size = 14,
             colour = "black")+
  draw_label("10", x = 0.31, y = 0.785, hjust = 0, vjust = 0,
             fontfamily = "", fontface = "plain", color = "black", size = 14,
             colour = "black")

```
